### 2019-11-25 Mathematical Statistics/ms_ch_03.rmd
mathematical statistics, ch3
mathematical statistics, ch3

p <- (sum(results[results >= observed]) + 1) / ( N + 1)
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..)) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Times, UA/AA vs Observed, p=", p ))

@@

p <- (sum(results[results <= observed]) + 1) / ( N + 1)
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..)) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Times, May/June, vs Observed, p=", round(p, 5)))

@@

>>>>

# two-sided test
p <- 2 * (sum(results[results >= observed]) + 1) / ( N + 1)
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Times, UA/AA vs Observed, p=", round(p, 5)),
        subtitle = paste0("Observed Value: ", round(observed, 4)))

@@

# two-sided test

p <- 2 * (sum(results[results <= observed]) + 1) / ( N + 1)
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Times, May/June, vs Observed, p=", round(p, 5)),
        subtitle = paste0("Observed Value: ", round(observed, 4)))

@@

### 3.6

In the flight delays case study in Section 1.1, the data contains flight delays for two airlines, American and United.

a.) Compute the proportion of times that each carrier's flight was delays more than 20 min.

Flights[, .(Delay20 = sum(Delay > 20) / .N), by = Carrier]

observed <- as.numeric(Flights[Carrier == "UA", .(Delay = sum(Delay > 20)/.N)] - Flights[Carrier == "AA", .(Delay = sum(Delay > 20)/.N)])

for(i in 1:N)
{
   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
   results[i] <- as.numeric(Flights[index, .(Delay = sum(Delay > 20)/.N)] - Flights[-index, .(Delay = sum(Delay > 20)/.N)])
}

p <- 2 * (sum(results[results >= observed]) + 1) / ( N + 1)
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Times, Over 20 Minutes, vs Observed, p=", round(p, 5)), 
        subtitle = paste0("Observed Value: ", round(observed, 4)))

+ Conduct a two-sided test to see if the difference in these proportions is statistically significant.

__Answer__: There is statistical significance with a P-value < 0.0001.

b.) Compute the variance in the flight delay lengths for each carrier.

Flights[, .(Variance = var(Delay)), by = Carrier]

observed <- var(Flights[Carrier == "UA"]$Delay) - var(Flights[Carrier == "AA"]$Delay)

for(i in 1:N)
{
   index <- sample(nrow(Flights), nrow(Flights[Carrier == "UA"]), replace = F)
   results[i] <- var(Flights[index]$Delay) - var(Flights[-index]$Delay)
}

p <- min(1, 2 * (sum(results[results >= observed]) + 1) / ( N + 1))
v <- p*(1 - p) / ( N + 1 )

ggplot(data.table(results)) +
   geom_histogram(aes(results, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed, col = "darkorange", linetype = 3, lwd = 1.2) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Flight Delay Time Variance, vs Observed, p=", round(p, 5)), 
        subtitle = paste0("Observed Value: ", round(observed, 4)))

+ Conduct a test to see if the variance for United Airlines differes from that of American Airlines.

__Answer__: There does not appear to be a statistically significant difference in the variance in delay times between airlines.

### 3.7

In the flight delays case study in Section 1.1, repeat Excercise 3.5 part __(a)__ using three test statistics,

+ i.) The mean of the United Airline delay times
+ ii.) The sum of the United Airline delay times
+ iii.) The difference in the means

Compare the P-values.

Make sure all three test statistics are computed within the same __for__ loop.

What do you observe?

```{r, echo = T, fig.height=7.5, fig.width=8}

ua.delay <- Flights[Carrier == "UA"]$Delay
aa.delay <- Flights[Carrier == "AA"]$Delay

observed.mean <- mean(ua.delay)
observed.sum <- sum(ua.delay)
observed.diff <- mean(ua.delay) - mean(aa.delay)

N <- 10e2 - 1
results <- matrix(nrow = N, ncol = 3)

for(i in 1:N)
{
   index <- sample(nrow(Flights), length(ua.delay), replace = F)

   results[i, 1] <- mean(Flights[index]$Delay)
   results[i, 2] <- sum(Flights[index]$Delay)
   results[i, 3] <- mean(Flights[index]$Delay) - mean(Flights[-index]$Delay)
}

dt.results <- data.table(results)
colnames(dt.results) <- c("Mean", "Sum", "MeanDiff")

p.mean <- ( sum( dt.results$Mean >= observed.mean ) + 1 ) / ( N + 1)
p.sum <- ( sum(dt.results$Sum >= observed.sum) + 1) / ( N + 1 )
p.diff <- ( sum(dt.results$MeanDiff >= observed.diff) + 1) / ( N + 1)

p1 <- ggplot(dt.results) +
   geom_histogram(aes(Mean, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed.mean, col = "darkorange", lwd = 1.2, linetype = 3) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("United Airlines - Mean Delay Time vs Observed, p=", round(p.mean, 5)), 
        subtitle = paste0("Observed Value: ", round(observed.mean, 4)))

p2 <- ggplot(dt.results) +
   geom_histogram(aes(Sum, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed.sum, col = "darkorange", lwd = 1.2, linetype = 3) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("United Airlines - Sum Delay Time vs Observed, p=", round(p.sum, 5)), 
        subtitle = paste0("Observed Value: ", round(observed.sum, 4)))

p3 <- ggplot(dt.results) +
   geom_histogram(aes(MeanDiff, fill = ..count..), bins = 30) +
   geom_vline(xintercept = observed.diff, col = "darkorange", lwd = 1.2, linetype = 3) +
   scale_y_continuous(labels = comma) +
   labs(title = paste0("Mean Delay Time Difference vs Observed, p=", round(p.diff, 5)), 
        subtitle = paste0("Observed Value: ", round(observed.diff, 4)))

grid.arrange(p1, p2, p3, nrow = 3)
